% Author:
% Date: January 21, 2026

\subsection{Constructing the Knowledge Base}
This section details the end-to-end pipeline for constructing a graph-oriented knowledge base derived from the WHO SMART Guidelines on HIV and Immunization. The process begins with the systematic extraction of domain-specific content from the source guidelines, capturing clinical recommendations and definitions. This raw data undergoes a rigorous parsing and normalization phase to transform unstructured text into standardized data nodes. Finally, these nodes are ingested into a Neo4j graph database, establishing the lateral relationships necessary to support complex querying and decision-support logic.

\subsubsection{Data Preprocessing}
The preprocessing phase serves as the entry point for the knowledge base construction, ensuring that structured data from the source guidelines is accurately prepared for downstream semantic analysis.

The workflow initiates with the JSON Extraction module, which ingests the WHO SMART Guidelines artifacts obtained from FHIR (Fast Healthcare Interoperability Resources) servers. This step isolates the relevant textual payloads, stripping away syntactic noise and structural metadata, to produce a clean stream of raw text.
\begin{verbatim}
Code: python/info_extraction/EntityExtraction.py
Function: extract_text_from_fhir
\end{verbatim}

Once isolated, this raw text is fed into Entity Extraction pipelines. There are two implementations:
- NER Named Entity Recognition (NER) pipeline
- General Purpose LLM extraction with Tool Universe

NER Named Entity Recognition (NER) pipeline
Named Entity Recognition (NER) pipeline. Unlike generic NLP methods, this approach can leverage a model that has been fine-tuned specifically for the biomedical domain to identify and tag medically relevant entities such as clinical terms, intervention protocols, and dosage parameters. This targeted extraction transforms the unstructured raw text into a set of annotated entities, ready for normalization. We've used the model \textit{blaze999/Medical-NER} to select the diseases that are within the content.
\begin{verbatim}
Code: python/info_extraction/EntityExtraction.py
Functions: extract_entities, enitity_cleaner_llm, entity_mapping
\end{verbatim}


After extracting the disease names, an LLM agent is prompted to select attributes related to each disease: diagnosis, management, symptoms, prognosis and contraindications. This agent has access to the entire knowledge base.
\begin{verbatim}
Code: python/info_extraction/EntityExtraction.py
Functions: extract_info_RAG
\end{verbatim}


\subsubsection{Knowledge Graph}
Neo4j Schema:
\begin{verbatim}
(:Disease)-[:HAS_DIAGNOSIS]->(:Diagnosis)
(:Disease)-[:HAS_MANAGEMENT]->(:Management)
(:Disease)-[:HAS_SYMPTOM]->(:Symptom)
(:Disease)-[:HAS_PROGNOSIS]->(:Prognosis)
(:Disease)-[:HAS_CONTRAINDICATION]->(:Contraindication)
\end{verbatim}


\subsubsection{LLM-Based Parsing with ToolUniverse}

In addition to the NER-based extraction pipeline, we implemented and validated a general-purpose LLM-based parsing workflow using \textit{ToolUniverse}, backed by the Gemini 2.5 Pro large language model. This workflow was designed to operate on both FHIR and non-FHIR clinical content and to produce structured JSON outputs suitable for downstream normalization and graph ingestion.

The ToolUniverse-based parsing logic is implemented as standalone scripts located under:
\begin{verbatim}
python/tool_universe/
\end{verbatim}


\paragraph{Non-FHIR Element Parsing}
Non-FHIR document parsing is implemented in:
\begin{verbatim}
python/tool_universe/tooluniverse_parse_non_fhir.py
\end{verbatim}

More details of the non-FHIR parsing are in section \ref{subsubsec:preprocessing_extensibility}: \nameref{subsubsec:preprocessing_extensibility}.

\paragraph{FHIR Element Parsing}
FHIR-based parsing is implemented in:
\begin{verbatim}
Code: python/tool_universe/tooluniverse_fhir.py
\end{verbatim}

This pipeline consumes index files generated by an upstream FHIR extraction function, which preprocesses WHO SMART Guidelines artifacts into indexed representations. These index files serve as inputs to the ToolUniverse-based FHIR parsing script, which extracts disease-level information and produces structured JSON outputs consistent with those generated from non-FHIR sources.

This approach enables FHIR and non-FHIR content to converge into a unified intermediate representation despite differences in their upstream extraction mechanisms.

\subsubsection{Lineage Generation and Provenance Tracking}

Both the FHIR and non-FHIR parsing pipelines generate lineage metadata as part of the extraction process. Lineage generation is implemented using retrieval context and document metadata exposed through ToolUniverse.

Each extracted disease attribute is linked back to:
\begin{itemize}
    \item Source document identifiers (file path)
    \item Retrieved text chunks (text chunk index on the original file, as well as starting and ending characters of the chunk)
    \item Query context used during extraction
\end{itemize}

This lineage information is preserved alongside the parsed JSON outputs and later propagated into the knowledge graph, enabling traceability and auditability of extracted facts.

The output of this process is a python dictionary saved as JSON in the format:

\begin{figure}[H]
\caption{JSON File Hierarchy}
\dirtree{%
.1 /.
.2 [Disease Name].
.3 importance \dots{} \textit{float}.
.3 symptoms \dots{} \textit{list[string]}.
.3 management \dots{} \textit{list[string]}.
.3 diagnosis \dots{} \textit{list[string]}.
.3 \_lineage.
.4 [Category Name].
.5 [Item Name].
.6 source\_path.
.6 resource\_id.
.6 score.
}
\end{figure}

\subsubsection{Neo4j Ingestion with Lineage}

Graph ingestion is implemented as a standalone Neo4j ingestion script located at:
\begin{verbatim}
Code: python/tool_universe/ingest_disease_normalized_with_lineage.py
\end{verbatim}

This script loads the JSON created on the previous step, then normalizes and ingests diseases into Neo4j, creating disease nodes and associated clinical attribute nodes, establishing typed relationships according to the graph schema. In addition, lineage metadata produced during parsing is persisted alongside graph entities, allowing each node and relationship to be traced back to its source documents and extraction context.

The ingestion process supports local execution and incremental updates, facilitating iterative validation and extension of the knowledge base.

